Paper Analysis

### 2025-04-29 Paper_WriteUp/Paper_Plots.qmd
adding chunk to plot RDs and volatilities for reviewers

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(scales)
library(ggridges)
library(here)
library(ggdist)
library(viridis)
library(PlayerRatings)
library(ggrepel)
library(ggpubr)
library(rstanarm)
library(ggh4x)
library(latex2exp)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

@@

init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

@@

### SET FIRST RD, VOLATILITY TO TRUE RD/VOLATILITY WE GET AFTER FINDING RDS/VOLATILITY IN RAW DATASET (IN INIT CALCULATION)

@@

#distribution of x_{ij} - for response to reviewer comment
```{r}
#import GM data
data_path = paste0(all_data_path[1], "/")

@@

>>>>

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(scales)
library(ggridges)
library(here)
library(ggdist)
library(viridis)
library(PlayerRatings)
library(ggrepel)
library(ggpubr)
library(rstanarm)
library(ggh4x)
library(latex2exp)
library(furrr)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

@@

init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

@@

@@

#distribution of x_{ij} - for response to reviewer comment
```{r}
#import data
data_path = paste0(all_data_path[1], "/")

@@

#plotting RDs and volatilities for reviewer response
```{r}
#import data
data_path = paste0(all_data_path[1], "/")

start = Sys.time()
init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

#get RDs
init_RDs = future_map_dfr(users, get_RDs, init_data) |>
  as_tibble()

end = Sys.time()
end - start #takes like 20 minutes

#RDs
RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3) +
  facet_wrap(~focal_user)

#save
ggsave(plot = RD_plot,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "RDs.png"), dpi = 1000,
       height = 6, width = 8)

#volatilities
vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3) +
  facet_wrap(~focal_user)

#save
ggsave(plot = vol_plot,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "volatilities.png"), dpi = 1000,
       height = 6, width = 8)
```
